A Transdiciplinary approach to developing a 3D spectral numerical coastal-flow model based on passive macroalgal rafting phenenomea to shed light on ambient and stochastic dispersal and connectivity of nearshore and offshore marine communities  around the coast of South Africa

### 2019-09-10 literature_review.Rmd
added section on kelp rafting. more to come. updated outputs. removed word template

---
title: A Transdiciplinary approach to developing a spectral numerical coastal-flow
  model based on passive macroalgal rafting phenenomea to shed light on ambient and
  stochastic dispersal and connectivity of nearshore and offshore marine communities
  and 3D hyperspectral hydrodynamic modelling in and around the coast of South Africa
author: "Ross Coppin"
output:
  pdf_document:
    toc: yes
  word_document:
    fig_caption: yes
    reference_docx: frontiers_template.docx
    toc: yes
---

@@

## Substrata

## Salinity

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment, and macroalgae are not exception. Wave exposure has been shown to play a role in determining distribution, abundance, diversity, composition, growth (Cousens 1982) and productivity (Pedersen and Nejrup 2012) of macroalgae communities. For example, the width, vertical zonation and diversity of algal communities often change predictably along gradients of wave exposure. Wave exposure may also drive macroalgae communities indirectly through the alteration in effect of another environmental driver. For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundary layer around the frond (???). The most important direct effect of wave exposure on macroalgal communities is through mechanical dislodgement, which ultimately leads to expiration. Wave exposure is a complex abiotic variable which varies spatially and temporarily in the marine environment. Furthermore, the degree to which a macroalgae community is exposed, is dependent on local site characteristics, such as bathymetry and local wind patterns. Despite this fact, macroalgae have been able to persist in often harsh and variable wave environments. Macroalgae are sessile organisms and incapabable of migrating when local conditions become unsuitable. Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological adaptation. The morphology of macroalgae are not fixed genetic traits. A study by Koehl et al. (2008) showed that transplanted Nereocystis luetkeana plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days. Advances in genetic techniques and taxonomy have revealed that species delineation based on morphology has been inaccurate, and organisms that were once considered two separate species are actually one species. For example, Moss (1948) investigated the anatomy, chemical composition of Fucus spiralis at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The authors found that individuals in exposed sites showed less branching of thalli as well as variation physiological components, such as organic nitrogen, mannitol, laminarian and alginic acid concentrations. The authors also noted a ‘crumpling effect’ displayed by individuals from exposed sites and inferred that this strategy may reduce overall drag. Other studies show that macroalgae in wave exposed environments have morphologies that reduce overall drag, increase strength of attachment or increase flexibility. There is also evidence that morphological adaptation is driven by currents, and in fact may be driving hydrological performance of macroalgae. Duggins et al. (2003) examined the direct and indirect flow effects on population dynamics, morphology and biomechanics of several understorey macroalgae species. These species included Costaria costata, Agarum fimbriatum, and Laminaria complanata and *Nereocystis luetkeana*. The results showed that in wave impacted sites (wave exposed) had higher rates of mortality, and no significance was found between survival of individuals and tidal or current velocity. The authors concluded that although tidal and current velocity did not play a significant role in determining kelp survival, it did play a role in morphological adaptation. The results from this study suggest that high current and tidal stresses are the main driver of kelp morphological adaptation. This in turn make those individuals more resilient to dislodgement to wave exposure. Wave exposure is stochastic in nature compared to tidal and ocean currents which are more regular in their frequency and magnitude. Therefore the regular forces of tidal and ocean currents may make kelp individuals more resilient to mechanical dislodgement over time.

@@

# Coastal flow

@@

@@

@@

An object will only float if the object density ($\rho$) is less than or equal to the density of the liquid in which the object is submerged, i.e. $\rho_{object} < \rho_{liquid}$. The bouyancy force (*BF*) is the force of the displaced liquid pushing the object in an upwards direction which is equal to the weight of the displaced liquid, $BF = W_{displaced-liquid}$. The weight of the displaced liquid can be calculated by multiplying the mass of the liquid by the gravitational constant $g$ and so the weight of the displaced liquid can be calculated by: $$W_{displaced-liquid} = mg$$ The equation is expanded upon by including the calculation for mass, $m = \rho V$, and so now the weight of the displaced liquid can be calculated as: $$mg = \rho_{liquid} V g$$ where $\rho$ represents the density of the liquid and $V$ represents the portion of the object submerged. Since the object is floating, the BF can also be represented by the weight of the object and as mentioned previously the weight of the displaced liquid can also be represented by $mg$, which in turn can be written as $W_{displaced-liquid} = W_{object}$. Now through substituting the calculations for weight, volume and the dimensions of the object, the equation can be simplied to: $$x= h\displaystyle\frac{\rho_o}{\rho_l}$$ where $h$ is the overall height of the object, $\rho_o$ represents the density of the object and $\rho_l$ represents the density of the liquid. It is important to note that this equation is only applicable to objects that are floating i.e. $\rho_{object} < \rho_{liquid}$. The portion of a floating object above the water surface can also be calcuated using a similar approach, and instead of calculating the submerged portion ($x$) the calculation will now solve for the fraction above the surface which is represented by $y$ (see figure \ref{fig:BF_1}). The same equation for calculating density ($\rho = \displaystyle\frac{m}{V}$) will also be used in solving for $y$. As mentioned previously, for floating objects the $BF = m_{object}g$ and the BF can also be defined as the weight of the displaced liquid ($BF = W_{displaced-liquid}$) and therefore the weight of the displaced liquid is equal to the weight of the object ($W_{displaced-liquid} = W_{object}$). On this basis equation ?? can be written as $$m_{liquid}g = m_{object}g$$
Substituting the equation for density into equation ?? the equation is now: $$\rho_{o}V_{o}g = \rho_{l}V_{l}g$$. Important to note is that $V_{o}$ represents the volume of displaced liquid i.e not the entire object volume (see figure \ref{fig:BF_1}). The constant $g$ is removed through cancellation and now the equation can be written as: $$\displaystyle\frac{V_l}{V_o} = \displaystyle\frac{\rho_o}{\rho_l}$$ We also now that the volume of the object submerged is equal to the volume of the displaced liquid and the fraction $\displaystyle\frac{V_l}{V_o}$ represents the portion of the object submerged. Therefore to calculate the portion above the surface this must be subtracted by 1, therefore: $$1-\displaystyle\frac{V_l}{V_o} = y$$
and since $\displaystyle\frac{V_l}{V_o} = \displaystyle\frac{\rho_o}{\rho_l}$ the equation can now be solved via density of the liquid and the object: $$y = 1 - \displaystyle\frac{\rho_o}{\rho_l}$$ Solving the equation algebraically it can now be written as:$$ y = \displaystyle\frac{\rho_l - \rho_o}{\rho_l}$$

@@

1. Determine if the hydrodynamic environment is the main driver of kelp morphology and if this is specific to a location

2. Simulate kelp rafting by means of a hydrodynamic modelling and calibrate this model with in situ beach-cast morphometric data.

3. Use the calibrated hydrodynamic model to investigate dispersal of microplastics along the west coast and south-west coast of South Africa.

>>>>

---
title: A Transdiciplinary approach to developing a 3D spectral numerical coastal-flow model based on passive macroalgal rafting phenenomea to shed light on ambient and stochastic dispersal and connectivity of nearshore and offshore marine communities  around the coast of South Africa
author: "Ross Coppin"
output:
  word_document:
    fig_caption: yes
    toc: yes
  pdf_document:
    toc: yes
bibliography: ref_lit_review.bib
---

@@

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment, and macroalgae are no exception. Wave exposure has been shown to play a role in determining distribution, abundance, diversity, composition, growth (Cousens 1982) and productivity (Pedersen and Nejrup 2012) of macroalgae communities. For example, the width, vertical zonation and diversity of algal communities often change predictably along gradients of wave exposure. Wave exposure may also drive macroalgae communities indirectly through the alteration in effect of another environmental driver. For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundary layer around the frond (???). The most important direct effect of wave exposure on macroalgal communities is through mechanical dislodgement, which ultimately leads to expiration. Wave exposure is a complex abiotic variable which varies spatially and temporarily in the marine environment. Furthermore, the degree to which a macroalgae community is exposed, is dependent on local site characteristics, such as bathymetry and local wind patterns. Despite this fact, macroalgae have been able to persist in often harsh and variable wave environments. Macroalgae are sessile organisms and incapabable of migrating when local conditions become unsuitable. Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological adaptation. The morphology of macroalgae are not fixed genetic traits. A study by Koehl et al. (2008) showed that transplanted Nereocystis luetkeana plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days. Advances in genetic techniques and taxonomy have revealed that species delineation based on morphology has been inaccurate, and organisms that were once considered two separate species are actually one species. For example, Moss (1948) investigated the anatomy, chemical composition of Fucus spiralis at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The authors found that individuals in exposed sites showed less branching of thalli as well as variation physiological components, such as organic nitrogen, mannitol, laminarian and alginic acid concentrations. The authors also noted a ‘crumpling effect’ displayed by individuals from exposed sites and inferred that this strategy may reduce overall drag. Other studies show that macroalgae in wave exposed environments have morphologies that reduce overall drag, increase strength of attachment or increase flexibility. There is also evidence that morphological adaptation is driven by currents, and in fact may be driving hydrological performance of macroalgae. Duggins et al. (2003) examined the direct and indirect flow effects on population dynamics, morphology and biomechanics of several understorey macroalgae species. These species included Costaria costata, Agarum fimbriatum, and Laminaria complanata and *Nereocystis luetkeana*. The results showed that in wave impacted sites (wave exposed) had higher rates of mortality, and no significance was found between survival of individuals and tidal or current velocity. The authors concluded that although tidal and current velocity did not play a significant role in determining kelp survival, it did play a role in morphological adaptation. The results from this study suggest that high current and tidal stresses are the main driver of kelp morphological adaptation. This in turn make those individuals more resilient to dislodgement to wave exposure. Wave exposure is stochastic in nature compared to tidal and ocean currents which are more regular in their frequency and magnitude. Therefore the regular forces of tidal and ocean currents may make kelp individuals more resilient to mechanical dislodgement over time.

@@

# Currents

@@

Dispersal and recruitment are recognised as important processes of marine ecosystems which are driven by a combination of wind and surface currents. This is particulary true for organisms which lack pelagic larvae, and are reliant on other modes of transport to reach new habitats through passive modes of dispersal from anthropogenic and natural sources. Anthropogenic sources include marine litter and structural debris while examples of organic sources include pumice and macroalgae in the form of kelp-rafts. Kelp-rafts have been identified as an important mode of passive dispersal of marine organisms, such as invertebrates [@helmuth1994; whichmann2012], epiphytes [@], grazers [@], invasive species [@] and macroalgae themselves [@] which either make landfall or are transported offshore where they eventually sink. A kelp-raft consists of an entanglement of macroalgae (one or multiple species) which have been dislodged from the benthic environment through hydrodynamic forces (mostly storms) and are positivley buoyant by means of gas-filled pneumatocysts and/or reproductive organs. Kelp-rafts are capable of travelling vast distances [@] and are considered important dispersal vectors in temperate latitudes, such as the Southern California Bight [@hobday2000] and the Northern Baltic Sea [@].

Kelp-raft abundance has been shown to vary temporally and spatially in the ocean and around coastlines. The temporal variability is dependent on seasonal growth patterns. For example, studies investigating the dispersal patterns of *Sargassum* in the West Pacific show increased abundance during growth seasons (spring and summer) as more biomass is available to fragment [@deysher1981; @kingsford1992]. Temporal variability may also be related to seasonal storm frequency where more storms in a particluar season relate to an increase in kelp-raft abundance. For example a study by @kingsford1995 investigated the contribution of *Macrosystis pyrifera* rafts to habitat conplexity in pelagic environments. The authors found that the abundance of *M. pyrifera* rafts increased during seasons where storms occured more frequently.

@@

An object will only float if the object density ($\rho$) is less than or equal to the density of the liquid in which the object is submerged, i.e. $\rho_{object} < \rho_{liquid}$. The bouyancy force (*BF*) is the force of the displaced liquid pushing the object in an upwards direction which is equal to the weight of the displaced liquid, $BF = W_{displaced-liquid}$. The weight of the displaced liquid can be calculated by multiplying the mass of the liquid by the gravitational constant $g$ and so the weight of the displaced liquid can be calculated by: $$W_{displaced-liquid} = mg$$ The equation is expanded upon by including the calculation for mass, $m = \rho V$, and so now the weight of the displaced liquid can be calculated as: $$mg = \rho_{liquid} V g$$ where $\rho$ represents the density of the liquid and $V$ represents the portion of the object submerged. Since the object is floating, the BF can also be expressed by the weight of the object and as mentioned previously the weight of the displaced liquid can also be expressed by $mg$, which in turn can be written as $W_{displaced-liquid} = W_{object}$. Now through substituting the calculations for weight, volume and the dimensions of the object, the equation can be simplied to: $$x= h\displaystyle\frac{\rho_o}{\rho_l}$$ where $h$ is the overall height of the object, $\rho_o$ represents the density of the object and $\rho_l$ represents the density of the liquid. It is important to note that this equation is only applicable to objects that are floating i.e. $\rho_{object} < \rho_{liquid}$. The portion of a floating object above the water surface can also be calcuated using a similar approach, and instead of calculating the submerged portion ($x$) the calculation will now solve for the fraction above the surface which is represented by $y$ (see figure \ref{fig:BF_1}). The same equation for calculating density ($\rho = \displaystyle\frac{m}{V}$) will also be used in solving for $y$. As mentioned previously, for floating objects the $BF = m_{object}g$ and the BF can also be defined as the weight of the displaced liquid ($BF = W_{displaced-liquid}$) and therefore the weight of the displaced liquid is equal to the weight of the object ($W_{displaced-liquid} = W_{object}$). On this basis equation ?? can be written as $$m_{liquid}g = m_{object}g$$
Substituting the equation for density into equation ?? the equation is now: $$\rho_{o}V_{o}g = \rho_{l}V_{l}g$$. Important to note is that $V_{o}$ represents the volume of displaced liquid i.e not the entire object volume (see figure \ref{fig:BF_1}). The constant $g$ is removed through cancellation and now the equation can be written as: $$\displaystyle\frac{V_l}{V_o} = \displaystyle\frac{\rho_o}{\rho_l}$$ We also now that the volume of the object submerged is equal to the volume of the displaced liquid and the fraction $\displaystyle\frac{V_l}{V_o}$ represents the portion of the object submerged. Therefore to calculate the portion above the surface this must be subtracted by 1, therefore: $$1-\displaystyle\frac{V_l}{V_o} = y$$
and since $\displaystyle\frac{V_l}{V_o} = \displaystyle\frac{\rho_o}{\rho_l}$ the equation can now be solved via density of the liquid and the object: $$y = 1 - \displaystyle\frac{\rho_o}{\rho_l}$$ Solving the equation algebraically it can now be written as:$$ y = \displaystyle\frac{\rho_l - \rho_o}{\rho_l}$$

Once the fractions of submerged and non-submerged portions of the object have been calculated the next calculation relevant to this study would be to calculate the weight needed to cause the object to sink or as least be completely submerged. This will relate to epiphyte loading on kelp rafts in order to determine raft longevity. If a weight is placed on the object in the conceptual model that has enough weight to fully submerge (but not sink) the object then an addition bouyancy force is needed, which will be represented by $BF_2$ which is equal to the additional mass that has been added which can be expressed as $mg$. Based on the previous calculations and the laws of physics we know that: $$W_{object} = mg = BF_2 = W_{add-liq-displaced}$$

The weight of the additional liquid displaced can now be calculated by substituting the density and volume of the additional displaced liquid along with the gravitational constant $g$ which is expressed as: $$W_{AL} = \rho_{l}V_{AL}g$$
where the subscript $AL$ represents the components of *additional liquid displaced*. The $V$ can also represent the change in volume ($\Delta V$) and so the expression can be expanded to: $$W_{AL} = \rho(\Delta V) g$$} where $\Delta V$ represents the change in volume of the object, i.e $y$.

@@

@@

1. Determine if the hydrodynamic environment is the main driver of kelp morphological characteristics using a numerical model

2. Simulate kelp rafting by means of particle dispersion modelling.

3. Conduct field experiments using artificial rafts and *in situ* kelp to track the movement of kelp around the South African coastline using custom GPS trackers.

4. Use the experimental data to calibrate the model and investigate the role of storms in ocean dispersal patterns.
